---
title: "Tom Uses NFLFastR"

---
```{r,echo=F}
rm(list=ls())
setwd("C:\\Users\\Thomas\\Desktop\\data\\nfl\\TSD")
```

#Setup
Packages and year + week inputs
```{r,echo=F,message=F,warning=F}
library(nflfastR)
library(tidyverse)
library(furrr)
library(stringr)

##inputs:
this_year = 2021
last_week = 7
has_run = TRUE
```

Helper fxns
```{r}
tg_trunc_d <- function(n,digits){
  a <- trunc(n*10^digits)/10^digits
  return(a)
}
```


#Gather play-by-play Data
```{r, message=F}
#don't ask
#closeAllConnections()

#get list of game ids from chosen year
game_id_list <- 
  fast_scraper_schedules(this_year) %>%
  filter(week<=last_week) %>%
  #filter(away_score != "NA") %>%
  pull(game_id)


#pbp is our data frame
# parallel fix added, still takes a second
future::plan("multisession", gc=TRUE)
pbp <- build_nflfastR_pbp(game_id_list)

#clean up
#rm(game_id_list)
#rm(last_week)
#rm(this_year)
```

Check Table (optional)
```{r}
#head(pbp)

#output (helpful for finding summary columns)
write.csv(pbp, file="pbp.csv")
```

#Transform play-by-play to Weekly
Set parameters
```{r}
min_pass_att = 1
min_rush_att = 1
min_tgt      = 1
```


Passing Stats
```{r}
pa.data <- pbp %>%
  filter(play_type=="pass") %>%
  group_by(passer, posteam, game_id) %>%
  summarise(
    id = paste0(passer, game_id),
    gid = paste0(posteam, game_id),
    pa.yds = sum(passing_yards, na.rm=TRUE),
    pa.tds = sum(touchdown),
    wpa = tg_trunc_d(sum(air_wpa, na.rm=TRUE),3),
    epa = tg_trunc_d(sum(qb_epa),2),
    pa.cmp = sum(complete_pass, na.rm=TRUE),
    pa.att = sum(pass_attempt, na.rm=TRUE),
    pa.cmp_pct = tg_trunc_d(pa.cmp / pa.att,3),
    pa.cpo = tg_trunc_d(mean(cpoe, na.rm = TRUE),2),
    pa.sck = sum(sack),
    pa.int = sum(interception),
    pa.1st = sum(first_down),
    pa.air = sum(air_yards, na.rm = TRUE),
    pa.dpb = sum(qb_dropback),
    pa.prob_dpb = tg_trunc_d(mean(xpass, na.rm=TRUE),3),
    pa.dpb_oe = tg_trunc_d(mean(pass_oe, na.rm=TRUE),3),
    pa.frc = sum(fumble_forced),
    pa.lst = sum(fumble_lost)
  ) %>% 
  filter(pa.att >= min_pass_att) %>% 
  arrange(desc(pa.yds))
pa.data <- pa.data[!duplicated(pa.data),]
#head(pa.data, 25)
#write.csv(pa.data, file="pa.csv")
```

Rushing Stats
```{r}
ru.data <- pbp %>%
  filter(play_type=="run") %>%
  group_by(rusher, posteam, game_id) %>%
  summarise(
    id = paste0(rusher, game_id),
    gid = paste0(posteam, game_id),
    ru.yds = sum(rushing_yards, na.rm=TRUE),
    ru.tds = sum(touchdown),
    wpa = tg_trunc_d(sum(wpa, na.rm=TRUE),3),
    epa = tg_trunc_d(sum(epa),2),
    ru.att = sum(rush_attempt, na.rm=TRUE),
    ru.1st = sum(first_down),
    ru.20z = sum(yardline_100<=20),
    ru.10z = sum(yardline_100<=10),
    ru.frc = sum(fumble_forced),
    ru.lst = sum(fumble_lost)
  ) %>%
  filter(ru.att >= min_rush_att) %>% 
  arrange(desc(ru.yds))
ru.data <- ru.data[!duplicated(ru.data),]
#head(ru.data, 25)
#write.csv(ru.data, file="ru.csv")
```

Receiving Stats
```{r}
re.data <- pbp %>%
  filter(play_type=="pass") %>%
  group_by(receiver, posteam, game_id) %>%
  summarise(
    id = paste0(receiver,game_id),
    gid = paste0(posteam, game_id),
    re.yds = sum(passing_yards, na.rm=TRUE),
    re.tds = sum(touchdown),
    wpa = tg_trunc_d(sum(comp_yac_wpa, na.rm=TRUE),3),
    epa = tg_trunc_d(sum(comp_yac_epa),2),
    re.cat = sum(complete_pass, na.rm=TRUE), 
    re.tgt = sum(pass_attempt, na.rm=TRUE),
    re.cat_pct = tg_trunc_d(re.cat / re.tgt,3),
    re.cpo = tg_trunc_d(sum(cpoe, na.rm = TRUE) / re.cat,2),
    re.1st = sum(first_down),
    re.air = sum(air_yards, na.rm = TRUE),
    re.yac = sum(yards_after_catch, na.rm = TRUE),
    re.xyc = tg_trunc_d(sum(xyac_median_yardage, na.rm = TRUE),1),
    re.frc = sum(fumble_forced),
    re.lst = sum(fumble_lost)
  ) %>% 
  filter(re.tgt >= min_tgt) %>% 
  arrange(desc(re.yds))
re.data <- re.data[!duplicated(re.data),]
#head(re.data, 25)
#write.csv(re.data, file="re.csv")
```

Team Volume Stats
```{r}
tm.data <- pbp %>% 
  group_by(posteam, game_id) %>% 
  summarise(
    gid = paste0(posteam, game_id),
    tm.yds = sum(yards_gained, na.rm = TRUE),
    tm.tds = sum(touchdown, na.rm=TRUE),
    tm.tgt = sum(complete_pass, na.rm = TRUE),
    tm.rat = sum(rush_attempt, na.rm = TRUE),
    tm.pyd = sum(passing_yards, na.rm = TRUE),
    tm.air = sum(air_yards, na.rm = TRUE),
    tm.ryd = sum(rushing_yards, na.rm = TRUE),
    tm.20z = sum(yardline_100<=20 & play_type == "run", na.rm = TRUE),
    tm.10z = sum(yardline_100<=10 & play_type == "run", na.rm = TRUE)
  )
tm.data <- tm.data[!duplicated(tm.data),]
#head(tm.data,10)
```

#Data Cleaning
Create one Weekly stats dataframe
```{r}
#join team stats into each category (pass, rush, rec)
tm.pa <- pa.data %>% 
  full_join(tm.data, by="gid") %>% 
  na.omit()
tm.pa <- tm.pa[!duplicated(tm.pa),]

tm.ru <- ru.data %>% 
  full_join(tm.data, by="gid") %>% 
  na.omit()
tm.ru <- tm.ru[!duplicated(tm.ru),]

tm.re <- re.data %>% 
  full_join(tm.data, by="gid") %>% 
  na.omit()
tm.re <- tm.re[!duplicated(tm.re),]
#head(tm.pa)


#join categories into one dataset
#this can take a long time
wk.data <- tm.pa %>% 
  full_join(tm.ru, by="id")
wk.data <- wk.data %>% 
  full_join(tm.re, by="id")
#head(wk.data)
```

Remove wasteful columns and reformat data
```{r}
#organize player names, teams, gameid, wpa, epa, tm stats
player <- wk.data$passer
tm <- wk.data$posteam.x.x
game_id <- wk.data$game_id.x.x
wpa <- wk.data$wpa.x
epa <- wk.data$epa.x
tm.tot.yds <- wk.data$tm.yds.x
tm.tot.tds <- wk.data$tm.tds.x
tm.rushes  <- wk.data$tm.rat.x
tm.target  <- wk.data$tm.tgt.x
tm.pyards  <- wk.data$tm.pyd.x
tm.ayards  <- wk.data$tm.air.x
tm.ryards  <- wk.data$tm.ryd.x
tm.20zone  <- wk.data$tm.20z.x
tm.10zone  <- wk.data$tm.10z.x
#sum(is.na(player))
tm[is.na(player)]       <- wk.data$posteam.x.y[is.na(wk.data$passer)]
game_id[is.na(player)]  <- wk.data$game_id.x.y[is.na(wk.data$passer)]
wpa[is.na(player)]      <- wk.data$wpa.y[is.na(wk.data$passer)]
epa[is.na(player)]      <- wk.data$epa.y[is.na(wk.data$passer)]
tm.tot.yds[is.na(player)] <- wk.data$tm.yds.x[is.na(wk.data$passer)]
tm.tot.tds[is.na(player)] <- wk.data$tm.tds.x[is.na(wk.data$passer)]
tm.rushes[is.na(player)]  <- wk.data$tm.rat.x[is.na(wk.data$passer)]
tm.target[is.na(player)]  <- wk.data$tm.tgt.x[is.na(wk.data$passer)]
tm.pyards[is.na(player)]  <- wk.data$tm.pyd.x[is.na(wk.data$passer)]
tm.ayards[is.na(player)]  <- wk.data$tm.air.x[is.na(wk.data$passer)]
tm.ryards[is.na(player)]  <- wk.data$tm.ryd.x[is.na(wk.data$passer)]
tm.20zone[is.na(player)]  <- wk.data$tm.20z.x[is.na(wk.data$passer)]
tm.10zone[is.na(player)]  <- wk.data$tm.10z.x[is.na(wk.data$passer)]
#player goes last
player[is.na(player)]   <- wk.data$rusher[is.na(wk.data$passer)]

tm[is.na(player)]       <- wk.data$posteam.x[is.na(wk.data$passer)&is.na(wk.data$rusher)]
game_id[is.na(player)]  <- wk.data$game_id.x[is.na(wk.data$passer)&is.na(wk.data$rusher)]
wpa[is.na(player)]      <- wk.data$wpa[is.na(wk.data$passer)&is.na(wk.data$rusher)]
epa[is.na(player)]      <- wk.data$epa[is.na(wk.data$passer)&is.na(wk.data$rusher)]
tm.tot.yds[is.na(player)] <- wk.data$tm.yds[is.na(wk.data$passer)&is.na(wk.data$rusher)]
tm.tot.tds[is.na(player)] <- wk.data$tm.tds[is.na(wk.data$passer)&is.na(wk.data$rusher)]
tm.rushes[is.na(player)]  <- wk.data$tm.rat[is.na(wk.data$passer)&is.na(wk.data$rusher)]
tm.target[is.na(player)]  <- wk.data$tm.tgt[is.na(wk.data$passer)&is.na(wk.data$rusher)]
tm.pyards[is.na(player)]  <- wk.data$tm.pyd[is.na(wk.data$passer)&is.na(wk.data$rusher)]
tm.ayards[is.na(player)]  <- wk.data$tm.air[is.na(wk.data$passer)&is.na(wk.data$rusher)]
tm.ryards[is.na(player)]  <- wk.data$tm.ryd[is.na(wk.data$passer)&is.na(wk.data$rusher)]
tm.20zone[is.na(player)]  <- wk.data$tm.20z[is.na(wk.data$passer)&is.na(wk.data$rusher)]
tm.10zone[is.na(player)]  <- wk.data$tm.10z[is.na(wk.data$passer)&is.na(wk.data$rusher)]
#player goes last
player[is.na(player)]   <- wk.data$receiver[
                                          is.na(wk.data$passer) & 
                                          is.na(wk.data$rusher)
                                         ]

#add as columns and remove originals plus some junk from merging
d1 <- c("passer","rusher","receiver","posteam.x","posteam.x.x","posteam.x.y","posteam.y","posteam.y.x","posteam.y.y","game_id.y","game_id.x","game_id.y.y","game_id.x.y","game_id.x.x","game_id.y.x","wpa.x","wpa.y","wpa","epa.x","epa.y","epa","gid","gid.x","gid.y","tm.yds.x","tm.tds.x","tm.tgt.x","tm.rat.x","tm.pyd.x","tm.air.x","tm.ryd.x","tm.20z.x","tm.10z.x","tm.yds.y","tm.tds.y","tm.tgt.y","tm.rat.y","tm.pyd.y","tm.air.y","tm.ryd.y","tm.20z.y","tm.10z.y","tm.yds","tm.tds","tm.tgt","tm.rat","tm.pyd","tm.air","tm.ryd","tm.20z","tm.10z")

wk.data <- cbind(player,
                 tm,
                 game_id,
                 wpa,
                 epa,
                 wk.data[,!names(wk.data) %in% d1],
                 tm.tot.yds,
                 tm.tot.tds,
                 tm.rushes,
                 tm.target,
                 tm.pyards,
                 tm.ayards,
                 tm.ryards,
                 tm.20zone,
                 tm.10zone
                 )


#optional export
wk.data0<-wk.data
wk.data0[is.na(wk.data0)]=0

write.csv(wk.data0,"2021-wk-data.csv")

#dataset for fantasy
fantasy.cols <- c(
  "player",
  "tm",
  "game_id",
  "pa.yds",
  "pa.tds",
  "pa.int",
  #"pa.1st",
  "ru.yds",
  "ru.tds",
  "ru.lst",
  #"ru.att",
  #"ru.1st",
  "re.yds",
  "re.tds",
  "re.cat"
  #"re.1st"
  #add other fantasy columns here as needed
)
ff.data <- wk.data[,names(wk.data) %in% fantasy.cols]
#head(ff.data)
write.csv(ff.data,"2021-ff-data.csv")
```


Output as csvs (optional)
```{r}
# write.csv(tm.pa[!duplicated(tm.pa),], "tm-pa.csv")
# write.csv(tm.ru[!duplicated(tm.ru),], "tm-ru.csv")
# write.csv(tm.re[!duplicated(tm.re),], "tm-re.csv")
# write.csv(wk.data,"wk-data.csv")
# write.csv(tm.data[!duplicated(tm.data),],"tm-data.csv")
```

#Data Cleaning pt. II (under construction)
Create one weekly stats dataframe
```{r}
# pl.data <- pa.data %>% 
#   full_join(ru.data,by="id") %>% 
#   na.omit()
# pl.data <- pl.data %>% 
#   full_join(re.data,by="id") %>% 
#   na.omit()
# head(pl.data,10)
# wk.data2 <- pl.data %>% 
#   full_join(tm.data,by="gid") %>% 
#   na.omit()
# head(wk.data2,10)
dpcols1 <- c("posteam","desc","play_type","yards_gained","td_player_name","touchdown")
has_run = TRUE

if(!has_run){
  
  pbp <- read.csv("pbp.csv")
  last_week = max(pbp$week)
  wk.data0 <- read.csv("2021-wk-data.csv")
  
}

wk.df <- pbp[pbp$week==last_week,dpcols1]
dpdf <- wk.df[ (wk.df$yards_gained>=50 & 
                  wk.df$touchdown == 1) 
               | 
               wk.df$yards_gained==69,]

print(dpdf[order(-dpdf$yards_gained),])

dpcols2 <- c("player","game_id","pa.yds","pa.tds","ru.yds","ru.tds","re.yds","re.tds")
wkd <- wk.data0[substr(wk.data0$game_id,1,7)==paste0(this_year&"_"&format(this_week,)))]
achi <- (wkd$pa.tds+wkd$ru.tds+wkd$re.tds)>=5 | 
            (wkd$ru.tds+wkd$re.tds)>=3 |
            wkd$ru.yds>=150 |
            wkd$re.yds>=150 |
            (wkd$ru.yds+wkd$re.yds)>=250 |
            wkd$pa.yds>=500
dpdf <- wkd[achi,
            dpcols2]
print(dpdf)
```

